Deep Reinforcement Learning \\ Actor-Critic and Trust region-based

### 2020-12-28 Paper/Paper.tex
Policy Gradient method section finished

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

@@

\section{What is Reinforcement Learning?}
\lettrine[nindent=0em,lines=3]{W}e refer as Machine learning (ML) to the group of algorithms that improve (learn) automatically through experience. Among this algorithms, we could say that there are three main classes (which depend on the kind of experience we provide):
\begin{itemize}
  \item \textbf{Supervised Learning}: The experience is given in the form of input and output examples, and the goal is to learn a general rule that maps inputs to outputs.
  \item \textbf{Unsupervised Learning}: The experience is given in the form of data (no outputs provided) and the goal is to discover hidden patterns in data
  \item \textbf{Reinforcement Learning}: No experience (data) is provided, instead a dynamic \emph{environment} is provided and an \emph{agent} must learn how to interact with it in order to achieve a goal.
\end{itemize}

\noindent In RL an agent interacts with and environment (assumed to be unknown), which in return may provides the agent with a rewards (see figure \ref{fig:RL_interaction}). The agent should take actions so the cumulative rewards is maximized. Ultimately, the goal of RL is that the agent learns a good (optimal) strategy from experimental trials and a relative simple received feedback, so his future rewards is maximized. In reality, the scenario could be a bot playing a game to achieve high scores, or a robot trying to complete physical tasks with physical items.

@@

As we can see in figure \ref{fig:RL_interaction}, the agent interacts with the environment sequentially. Therefore, on time step $t \in \{1,...,T\}$, the agent observe the state of the environment $s_t \in \mathcal{S}$, acts in consequence executing the action $a_t \in \mathcal{A}$, receive its reward $r_t \in \mathcal{R}$ from the environment and observes the new state $s_{t+1} \in \mathcal{S}$.
The tuple $(s_t, a_t, r_t, s_{t+1})$ is known as \emph{transition} step.
This interactions continues until the agent: \textit{i}) completes the task successfully; \textit{ii}) fail or a time limit is reached. This chain of transition steps from the beginning ($t=1$) until the end ($t=T$) is call \emph{Episode}.
We are interested in environments for which the model that defines them is unknown (\emph{Model-free RL}), and has to be learned explicitly as part of the algorithm (otherwise the optimal solution can be found by means of \emph{Dynamec Programming} (DP)).

@@

However, note that in fact we are not interested in maximizing the reward on each time step, but the \emph{cumulative reward} in the long run instead. Also, we must think about time as a valuable resource, since in practices we would like to finish a task as soon as possible. This cumulative reward is called \emph{Return} and is defined as:
\begin{equation}
  G_t := \sum_{k=0}^T \gamma^k R_{t+k+1}
\end{equation}
where $\gamma \in [0,1]$ is the discount factor.

Recall that the ultimate goal of RL is to find an optimal strategy, such that it maximize the Return. However, how to choose the actions that will lead us to maximize the return? The answer is the \emph{Policy} $\pi$. The policy is the functions that defines the agent’s behavior and maps a state $s$ to an action $a$. this function can be either deterministic or stochastic:

@@

\noindent However, the Action-value function tell us how valuable and action is w.r.t. the long term (return). But, what if we only want to know the value that taking the action $a$, following the policy $\pi$, could bring in the short term (i.e. when it is executed in an environment to go from state $s$ to the next one $s'$)? This question is answered by the \textbf{Advantage} function:

@@

There are several ways to approximate an optimal strategy. For example, in \emph{action-value} methods we learn the values of actions and then select
actions based on their estimated action values (as it is done in \textit{Q-learning}). Nevertheless, this means that their policies would not even exist without the action-value estimates \cite{Sutton1998}.

However, in \emph{Policy Gradient} we consider methods that instead learn a parameterized policy (e.g. an artificial neural network) that can select actions without consulting a value function\footnote{A value function may still be used to learn the policy parameter, but is not required for action selection.}.
Therefore, we write
$$\pi(a|s,\boldsymbol{\theta}) = Pr(a_t=a | s_t=s, \boldsymbol{\theta}_t=\boldsymbol{\theta})$$
 to denote the probability that the action $a$ is taken, given the parameters $\boldsymbol{\theta}$ and that the environment is in state $s$ at time $t$.
Then, we can approximate an optimal policy by means of gradient ascent while maximizing a performance measure $J(\boldsymbol{\theta})$:

\begin{equation}
  \boldsymbol{\theta}_{t+1} = \boldsymbol{\theta}_{t} + \alpha \nabla J(\boldsymbol{\theta}_t)
\end{equation}

es importante poner el hat! busca como

Note that in practice $\nabla J(\boldsymbol{\theta}_t)$ is a stochastic estimate whose expectation approximates the gradient of the performance measure with respect to its argument $\boldsymbol{\theta}_t$.

Then, how to define the performance measure $J(\boldsymbol{\theta})$? Since we seek to maximize the return $G$, then we define it as the value of the first state of the episode\footnote{Note that this definition only applies to the episodic case (i.e. discrete time steps). The definition and explanation of the continuous case is out of the scope of this work.}, i.e. $J(\boldsymbol{\theta}):=v_{\pi_{\boldsymbol{\theta}}}(s_0)$

poner a que es igual es igual la v, luego a que es igual el gradiente de la v (meniona aque el teorema del gradiente ese y la pagina del barto para mayor referencia)
luego la regla para gradient acent con el nuevo gradiente (pag 327, 13.8)

luego explica por que el baseline y pon la nueva regla de gradient ascent (13.11, pag 329)

fin de seccion

@@

>>>>

%\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[margin=18mm,top=32mm,columnsep=20pt]{geometry}
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

@@

\section{What is Reinforcement Learning?}
\lettrine[nindent=0em,lines=3]{W}e refer as Machine learning (ML) to the group of algorithms that improve (learn) automatically through experience. Among this algorithms, we could say that there are three main classes (which depend on the kind of experience we provide):
\begin{itemize}
  \item \textbf{Supervised Learning}: The experience is given in the form of input and output examples, and the goal is to learn a general rule that maps inputs to outputs.
  \item \textbf{Unsupervised Learning}: The experience is given in the form of data (no outputs provided) and the goal is to discover hidden patterns in data.
  \item \textbf{Reinforcement Learning}: No experience (data) is given, instead a dynamic \emph{environment} is provided and an \emph{agent} must learn how to interact with it in order to achieve a goal.
\end{itemize}

\noindent The best way to summarize RL is ``learning by interaction''; An agent interacts with and environment, which in return provides the agent with feedback in form of a reward (see figure \ref{fig:RL_interaction}). Ultimately, the goal of RL is that the agent learns a good (optimal) strategy from experimental trials, and a relative simple received feedback, so that the cumulative rewards is maximized. In reality, the scenario could be a bot playing a game to achieve high scores, or a robot trying to complete physical tasks with physical items.

@@

As we can see in figure \ref{fig:RL_interaction}, the agent interacts with the environment sequentially. Therefore, on time step $t \in \{1,...,T\}$, the agent observe the state of the environment $s_t \in \mathcal{S}$, acts in consequence executing the action $a_t \in \mathcal{A}$, receive its reward $r_t \in \mathcal{R}$ from the environment and observes the new state $s_{t+1} \in \mathcal{S}$.
The tuple $(s_t, a_t, r_t, s_{t+1})$ is known as \emph{transition} step.
This interactions continues until the agent: \textit{i}) Completes the task successfully; \textit{ii}) Fails; \textit{iii}) Reach a time limit. This chain of transition steps from the beginning ($t=1$) until the end ($t=T$) is call \emph{Episode}.
We are interested in environments for which the model that defines them is unknown (\emph{Model-free RL}), and has to be learned explicitly as part of the algorithm (otherwise the optimal solution can be found by means of \emph{Dynamec Programming} (DP)).

@@

However, note that in fact we are not interested in maximizing the reward on each time step but the \emph{Return} instead, which is the future cumulative reward\footnote{This might sound counterintuitive. However, think of it as when you have to make less rewarding decisions in the short term (e.g. watching Netflix instead of studying) in order to achieve a greater long-term goal (e.g. approving an exam).}:
\begin{equation}
  G_t := \sum_{k=0}^T \gamma^k r_{t+k+1}
\end{equation}
where $\gamma \in [0,1]$ is the discount factor meant to encourage the agent to finish the task as soon as possible, since in practice time is a valuable resource.

Recall that the ultimate goal of RL is to find an optimal strategy, such that it maximize the Return. Then, how to choose actions that help us to achieve this? The answer is the \emph{Policy} $\pi$. The policy is the functions that defines the agent’s behavior and maps a state $s$ to an action $a$. This function can be either deterministic or stochastic:

@@

mover esto a action value section, ya que la v ahora va a tener parametros w

\noindent However, the Action-value function tell us how valuable and action is w.r.t. the long term (return). But, what if we only want to know the value that taking the action $a$ (following the policy $\pi$) could bring in the short term (i.e. when it is executed in an environment to go from state $s$ to the next one $s'$)? This question is answered by the \textbf{Advantage} function:

@@

There are several ways to approximate an optimal strategy. For example, in \emph{action-value} methods we learn the values of actions and then select
actions based on their estimated action values (e.g. \textit{Q-learning}). Nevertheless, this means that their policies would not even exist without the action-value estimates \cite{Sutton1998}.

However, in \emph{Policy Gradient} we consider methods that instead learn a parameterized policy (e.g. an artificial neural network) that can select actions without consulting a value function\footnote{A value function may still be used to learn the policy parameter, but is not required for action selection.}.
Therefore, we write
$$\pi(a|s,\boldsymbol{\theta}) = Pr(A_t=a | S_t=s; \boldsymbol{\theta}_t=\boldsymbol{\theta})$$
 to denote the probability that the action $a$ is taken, given the parameters $\boldsymbol{\theta}$ and that the environment is in state $s$ at time $t$.
Then, we can approximate an optimal policy by means of gradient ascent while maximizing a performance measure $J(\boldsymbol{\theta})$:

\begin{equation}
  \boldsymbol{\theta}_{t+1} = \boldsymbol{\theta}_{t} + \widehat{\alpha \nabla J(\boldsymbol{\theta}_t)}
  \label{eq:learn_rule}
\end{equation}

\noindent where $\widehat{\nabla J(\boldsymbol{\theta})}$ is a stochastic unbiased estimator of $\nabla J(\boldsymbol{\theta})$ (i.e. $\mathbb{E}[\widehat{\nabla J(\boldsymbol{\theta})}] = \nabla J(\boldsymbol{\theta})$).

Then, how to define the performance measure $J(\boldsymbol{\theta})$?
There are many ways to define $J$, which depend mainly on the method used to approximate the policy.
However, since we are considering only episodic environments\footnote{The definition and explanation of the continuous case is out of the scope of this work.}, then we seek to maximize the return in a given time step ($G_t$), and therefore we define it as the value of the state $s_t$ , i.e. $J(\boldsymbol{\theta}):=V_{\pi_{\boldsymbol{\theta}}}(s_t)$.

Since we are approximating an optimal policy, we need to express $\nabla J(\boldsymbol{\theta})$ in terms of $\pi(\boldsymbol{\theta})$. Therefore, after marginalize $V_{\pi_{\boldsymbol{\theta}}}(s_t)$ over the action space $\mathcal{A}$ and applying Baye's rule we have:

\begin{equation}
  \begin{split}
    \nabla J({\boldsymbol{\theta}}) &= \nabla V_{\pi_{\boldsymbol{\theta}}}(s_t)\\
    &= \nabla \sum_{a_t \in \mathcal{A}} \pi (a_t|s_t; \boldsymbol{\theta}) Q_{\pi_{\boldsymbol{\theta}}}(s_t,a_t)
  \end{split}
  \label{eq:grad}
\end{equation}

The problem is that the right hand side of equation \ref{eq:grad} depends on both the action selections and the distribution of states in which those selections are made, and that both of these are affected by the policy parameter. Fortunately, with the aid of the \emph{Policy Gradient Theorem} we can reformulate $\nabla J({\boldsymbol{\theta}})$ to be proportional to $\nabla \pi (a_t|s_t; \boldsymbol{\theta})$\footnote{A complete proof to this result can be found in page 325 of \cite{Sutton1998}.}:

\begin{equation*}
\resizebox{.95\linewidth}{!}{%
  $\begin{split}
    \nabla J({\boldsymbol{\theta}}) &\propto
    \sum_{s_t \in \mathcal{S}} \mu (s_t) \sum_{a_t \in \mathcal{A}} Q_{\pi_{\boldsymbol{\theta}}}(s_t,a_t) \nabla \pi (a_t|s_t; \boldsymbol{\theta}) \\
    &=\sum_{s_t \in \mathcal{S}} \mu (s_t) \sum_{a_t \in \mathcal{A}} \pi (a_t|s_t; \boldsymbol{\theta}) Q_{\pi_{\boldsymbol{\theta}}}(s_t,a_t) \frac{\nabla \pi (a_t|s_t; \boldsymbol{\theta})}{\pi (a_t|s_t; \boldsymbol{\theta})} \\
    &=\sum_{s_t \in \mathcal{S}} \mu (s_t) \sum_{a_t \in \mathcal{A}} \pi (a_t|s_t; \boldsymbol{\theta}) Q_{\pi_{\boldsymbol{\theta}}}(s_t,a_t) \nabla \log{\pi (a_t|s_t; \boldsymbol{\theta})} \\
    &= \mathbb{E}[Q_{\pi_{\boldsymbol{\theta}}}(S_t,A_t) \nabla \log{\pi (A_t|S_t; \boldsymbol{\theta})}] \\
    &= \mathbb{E}[G_t \nabla \log{\pi (A_t|S_t; \boldsymbol{\theta})}] \\
  \end{split}
  $}
\end{equation*}

\noindent where $\mathbb{E}$ is the expectation over the distribution of states $\mu(s_t)$ and actions $\pi (a_t|s_t; \boldsymbol{\theta})$; $A_t$ and $S_t$ are random variables ($A_t \sim \mu(s_t)$ and $S_t \sim \pi (a_t|s_t; \boldsymbol{\theta})$); and $Q_{\pi_{\boldsymbol{\theta}}}(S_t,A_t)$ can be replaced by $G_t$ since $\mathbb{E}[G_t| S_t,A_t] = Q_{\pi_{\boldsymbol{\theta}}}(S_t,A_t)$.

Therefore, our learning rule defined in \ref{eq:learn_rule} can be updated to:

\begin{equation*}
  \boldsymbol{\theta}_{t+1} = \boldsymbol{\theta}_{t} + \alpha G_t \nabla \log{\pi (A_t|S_t; \boldsymbol{\theta})}
\end{equation*}

However, the \emph{Policy Gradient Theorem} can be generalized to include a comparison of the action value to an arbitrary baseline $b(s)$ \cite{Sutton1998}, which can be any function, even a random variable, as long as it does not vary with a:

\begin{equation*}
  \nabla J({\boldsymbol{\theta}}) \propto
  \sum_{s_t \in \mathcal{S}} \mu (s_t) \sum_{a_t \in \mathcal{A}} \left( Q_{\pi_{\boldsymbol{\theta}}}(s_t,a_t) - b(s_t) \right) \nabla \pi (a_t|s_t; \boldsymbol{\theta})
\end{equation*}

the last equation remains valid because the subtracted quantity is zero:

\begin{equation*}
  \begin{split}
    \sum_{a_t \in \mathcal{A}} b(s_t) \nabla \pi (a_t|s_t; \boldsymbol{\theta}) &=
    b(s_t) \sum_{a_t \in \mathcal{A}} \nabla \pi (a_t|s_t; \boldsymbol{\theta})\\
    &= b(s_t) \nabla 1 \\
    &= 0
  \end{split}
\end{equation*}

Finally, we can use this to derive a new update rule that includes a general baseline:

\begin{equation}
  \boldsymbol{\theta}_{t+1} = \boldsymbol{\theta}_{t} + \alpha (G_t - b(S_t)) \nabla \log{\pi (A_t|S_t; \boldsymbol{\theta})}
  \label{eq:learn_rule_bl}
\end{equation}

@@

poner aqui la foto del video (el diagrama de ben)
